A comprehensive framework to capture the arcana of neuroimaging analysis

### 2018-10-13 manuscript/manuscript.tex
touched up manuscript listings

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{listings}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{color}

@@

\lstset{
    basicstyle=\footnotesize,
%  captionpos=b,                    % sets the caption-position to bottom
    language=Python,                 % the language of the code
    showstringspaces=false,          % underline spaces within strings only
    stringstyle=\color{stringcolour},     % string literal style
    commentstyle=\color{commentcolour},    % comment style
    keywordstyle=\color{keywordcolour}\bfseries,       % keyword style
}

@@

\author{
	Thomas G. Close\affmark[1,2]\and
	Phillip G. D. Ward\affmark[1,3,4] \and
	Francesco Sforazzini\affmark[1] \and
	Wojtek Goscinski\affmark[6] \and
	Zhaolin Chen\affmark[1,5] \and
	Gary F. Egan\affmark[1,3,4]
}

@@

Despite the availability of well-established neuroimaging analysis
packages \cite{cox_afni:_1996,smith_advances_2004,friston_statistical_2007,tournier_mrtrix:_2012}, the arcana of neuroimaging analysis is substantial due to
the range of available tools, tuneable parameters, and imaging sequences
involved \cite{cusack_automatic_2015}. The distribution of complete
``end-to-end'' workflows, from acquired data to publication results, is
necessary for routine reproduction because of the effort required to
accurately exactly reimplement such analyses \cite{kennedy_neuroimaging_2018}. It is also
difficult to adapt existing workflows to new studies without detailed
knowledge of their design. Therefore, flexible, portable and complete
workflows are important to promote reproduction and code reuse in
neuroimaging research.

A barrier to designing portable and complete workflows is the
heterogeneity of data storage conventions \cite{marcus_extensible_2007,das_loris:_2012,gorgolewski_brain_2016}.
To address this, the emerging Brain
Imaging Data Standard (BIDS) \cite{gorgolewski_brain_2016} provides a way
to standardise the storage of neuroimaging data in file system
directories. BIDS specifies strict file and directory naming
conventions, which facilitate the design of portable \emph{BIDS Apps}
(https://bids-apps.neuroimaging.io). However, for research groups with
sufficient informatics support, software-managed repositories \cite{marcus_extensible_2007,das_loris:_2012} can provide additional features, such as
flexible access-control and automated pipelines. Therefore, for
published workflows, the choice of repository should be transparent in
order to maximise their audience.

While neuroimaging analyses are generally amenable to standardisation
\cite{kennedy_neuroimaging_2018}, minor modifications are often required to accommodate
idiosyncrasies of the acquisition protocols in use at different sites
\cite{esteban_fmriprep:_2018}. Therefore, workflows may require conditional
logic in construction or execution to be portable. Nipype is a flexible
Python framework for neuroimaging analysis in which workflows are
constructed programmatically in Python \cite{gorgolewski_nipype:_2011}.
Programmatic construction allows for rich control-flow logic that is not
readily available in alternative workflow frameworks \cite{cusack_automatic_2015,achterberg_fastr:_2016,amstutz_common_2016}, and has been used
to implement workflows that are robust to differences in fMRI protocols
across a large number of sites \cite{esteban_fmriprep:_2018}.

The trend towards large multi-site and multi-contrast datasets collected
over a number of years \cite{van_essen_human_2012,thompson_enigma_2014,sudlow_uk_2015}
presents additional challenges to workflow design.
Analysis packages are constantly being developed and improved, so the
state-of-the-art workflow for a particular analysis can change over
time. Therefore, it is challenging to ensure workflows are applied
consistently over the course of long studies \cite{cusack_automatic_2015}.

While analysis workflows for different contrasts and modalities are
typically implemented independently, they can share common processing
steps (e.g. non-linear registration to standard space, surface
parcellation) and their outputs may need to be integrated to produce
publication results. For large scale studies, which are typically
processed on the cloud or high-performance computing (HPC) clusters,
rerunning common segments can lead to significant increases in
computation time and therefore project cost. In addition, duplication of
processing segments increases time for manual quality control (QC),
making the reuse of intermediate derivatives a practical requirement for
some large studies \cite{schreiber_using_2018}.

@@

The utility of the Arcana framework is demonstrated by the
implementation of analysis suites for T1, T2* and diffusion weighted MRI
(dMRI) data and the application of dMRI tractogram \cite{tournier_mrtrix:_2012} and
composite vein image \cite{ward_combining_2018} workflows to data collected from a
healthy subject.

@@

\begin{figure}
\lstinputlisting[language=Python, firstline=2, lastline=57]{listings/example_study.py}
\caption{Example data and parameter specifications. The data
specification specifies two ``acquired'' file sets, `one' and `ten' and
ten derived file sets that can be derived from them, at least
indirectly. Each derived data spec, specifies the name of the pipeline
constructor that creates the pipeline that derives them. Parameter
specifications specify a name and default value for free parameters of
the Study class}
\label{list:data_spec}
\end{figure}

@@

Similar to data specifications, parameter specifications are included in
the Study class by providing a list of \emph{ParameterSpec} objects to
the \emph{add\_parameter\_specs} class attribute. ParameterSpec objects
are initialised with a name and default value. Special parameters that
specify a qualitative change in the analysis, for example using ANTs
registration \cite{avants_reproducible_2011} instead of FSL registration \cite{smith_advances_2004}, are specified by the \emph{SwitchSpec} subtype.
SwitchSpecs take a name, default value and a list of accepted values at
initialisation.

@@

\begin{figure}
\lstinputlisting[language=Python, firstline=58, lastline=108]{listings/example_study.py}
\caption{Example pipeline constructor method. Pipelines are
created using the \emph{create\_pipeline} method of the Study class.
Pipeline objects are thin wrappers around Nipype workflows to allow for
namespace management}
\label{list:pipeline_constructor}
\end{figure}

@@

Any external software packages required by a node should be referenced
in the \emph{requirements} keyword argument as a list of
\emph{Requirement} objects when the node added to the pipeline. These
requirements can be used by the Processor to configure environment
modules \cite{furlani_modules:_1991} if present, to ensure the required software
packages are available. Similarly, the expected memory requirements and
wall time for the node execution should be provided as keyword args,
\emph{memory}, and \emph{wall\_time}, so they can be inserted into job
scripts submitted to schedulers by HPC Processors.

@@

Instead of setting the class attributes directly, the data and parameter
specifications are set by the metaclas\emph{s}, in order to combine with
specifications from base classes. The combined data and parameter
specifications are constructed by visiting the class' bases in reverse
method resolution order (MRO) and adding specifications from their
\emph{add\_data\_specs}, \emph{add\_parameter\_specs} attributes,
overriding previously added specifications with matching names. Note
that in this scheme, specifications can only be appended or overridden
but not removed by Study subclasses so as not to break workflows
inherited from base classes.

@@

All MultiStudy classes must inherit from the Multi\emph{Study} base
class and be created by the Multi\emph{StudyMetaClass} metaclass or
subclasses thereof. As in the case of subclassing the standard Study
class, additional data and parameter specifications can be added to the
class via \emph{add\_data\_specs} and \emph{add\_parameter\_specs}
respectively for additional analysis not included in the sub-studies.

@@

In Arcana, repository access is encapsulated within modular
\emph{Repository} objects to enable switching between different
repositories and repository types at analysis ``application time''
(Figure 2). There are currently three supported repository types, XNAT
\cite{marcus_extensible_2007}, BIDS \cite{gorgolewski_brain_2016} and a simple
``directory'' format, which are encapsulated by \emph{XnatRepository},
\emph{BidsRepository}, and \emph{DirectoryRepository} classes
respectively.

@@

\begin{table}
\caption{Storage locations of derived data for each repository
type. Derivatives are stored in separate namespaces for each Study
instance to enable multiple analyses on the same datasets with different
parameterisations. Where `...' is the location of the
directory or MrSession that holds the derivatives, \emph{subj} = subject ID, \emph{vis} = visit ID,
\emph{study} = name of the Study instance, and \emph{pl} = name
of pipeline.}
\label{tbl:repo-structure} 
\begin{tabular}{llll}
\hline\noalign{\smallskip}
\textbf{Type} & Plain-directory & BIDS & XNAT  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\textbf{Derivatives} & /\emph{subj}/\emph{vis}/\emph{study} & /derivatives/\emph{study}/\emph{subj}/\emph{vis} & /\emph{subj}/\emph{vis}\_\emph{study}\\
\textbf{Fields} & \D/\_\_fields\_\_.json & \D/\_\_fields\_\_.json & MrSession XML \\ 
\textbf{Provenance} & \D/\_\_prov\_\_/\emph{pl}.json & \D/\_\_prov\_\_/\emph{pl}.json & \D/\_\_prov\_\_/\emph{pl}.json \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

@@

\begin{table}
\caption{Abstract methods in the base Repository class that
need to be implemented by platform-specific sub-classes.}
\label{tbl:abstract-repo-methods} 
\begin{tabular}{ll}
\hline\noalign{\smallskip}
\textbf{Method name} & Function \\
\noalign{\smallskip}\hline\noalign{\smallskip}
find\_data() & Queries repository to find all existing filesets and
fields \\
get\_fileset(Fileset) & Caches file-set (if necessary) and returns the file path. \\
get\_field(Field) & Retrieves and returns the value of the field from
the repository \\
put\_fileset(Fileset) & Inserts file-set into the repository and updates cache. \\
put\_field(Field) & Inserts the value of the Field into the repository \\
connect() & Opens a connection to the repository (optional). \\
disconnect() & Closes connection to repository (optional)\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

@@

Software versions are managed by the processor if environment modules
\cite{furlani_modules:_1991} are installed on the system. These are loaded and
unloaded before the interface of a workflow node is run based on the
requirements specified for the node during construction of the pipeline
(see \emph{Pipeline constructors}. Maps from requirement names to
module names in the processing environment can be passed to the
Processor at initialisation. For plugins that submit jobs to HPC
schedulers (e.g. SLURM, SGE), node memory and wall-time requirements are
inserted into jobscripts where required.

@@

The Arcana framework is distributed as a publicly available software
package via GitHub
(\href{http://github.com/MonashBI/arcana}{{http://github.com/MonashBI/arcana}})
and the Python Package Index (PyPI)
(\href{https://pypi.org/project/arcana/}{{http://pypi.org/project/arcana/}}).
Study classes for T1, T2* and diffusion weighted MRI data have been
implemented as part of the \emph{Biomedical imAgiNg ANAlysis (Banana)}
package
(\href{https://github.com/MonashBI/banana}{{http://github.com/MonashBI/banana}};
\href{http://pypi.org./project/banana}{{http://pypi.org./project/banana}}).
All three classes, \emph{T1wStudy}, \emph{T2starwStudy} and
\emph{DmriStudy}, inherit generic image analysis methods, such as
registration and brain extraction, from the base class \emph{MRIStudy}.
The T1Study and T2starStudy classes are aggregated into the
\emph{T2starwT1wStudy} MultiStudy class to combine analysis of T2* and
T1-weighted contrasts.

The DmriStudy class implements diffusion tensor metrics, FA and ADC, as
well as global tractography using streamlines tracking from the MRtrix
toolbox \cite{tournier_improved_2010,tournier_mrtrix:_2012} (Figure 9).

@@

The \emph{T2starwT1wStudy} class implements an algorithm to generate
\emph{composite vein images} from the combination of QSM and SWI
contrasts derived from the T2*-weighted acquisition with an vein atlas
registered to the T1-weighted acquisition \cite{ward_combining_2018} (Figure 10).

@@

To test the analysis workflows implemented in ArcanaPaper, a healthy
volunteer was scanned using a 3T Siemens Skyra with a 32-channel head
and neck coil. The protocol was a T1-weighted MPRAGE (1mm contiguous,
matrix size 256x240x192, FOV 256x240x192, TE = 2.13ms, TR, = 2300ms, TI
= 900ms, bandwidth=230Hz/pixel), GRE (1.8 mm contiguous, matrix size
256x232x72, FOV 230x208x130, TE = 20ms, TR, = 30ms,
bandwidth=120Hz/pixel), diffusion MRI (1.2 mm contiguous, matrix size
110x100x60, FOV 256x240x192, TE = 95ms, TR, = 8200ms, 33 diffusion
directions with b=1500 mm2/s and 3 b=0, bandwidth=781Hz/pixel)
(\textbf{cite Lifespan paper ??}). Methods for Figure 11, 12, and 13 in
the ArcanaPaper class were applied to the acquired dataset to using the
\emph{SlurmProcessor} to submit jobs to the the MASSIVE HPC cluster
\cite{goscinski_multi-modal_2014} in order to generate the figures.

Figure \ref{fig:veins} displays composite vein images and vein masks for the healthy
volunteer along with the SWI and QSM intermediate derivatives and the
input vein atlas. The derived vein images are comparable to those
generated by the original implementation \cite{ward_combining_2018}. The complete
workflow took ?? min to complete over ?? compute cores.

@@

Software frameworks \cite{yacoub_pattern-oriented_2004} have been successful in
improving code quality and efficiency of development in a variety of
contexts \cite{moore_professional_2008,white_hadoop:_2012,abadi_tensorflow:_2016}. By
factoring out common elements, only features that are specific to the
given application need to be implemented by the analysis designer, and
the common elements become battle hardened through repeated use. Arcana
handles many of the menial tasks involved with workflow implementation,
such as data retrieval and storage, format conversions, and provenance,
reducing the time and effort required to implement robust workflows from
acquired data to publication results.

An oft-repeated mantra in the open-source software movement dubbed
Linus' Law is that "given a large enough beta-tester and co-developer
base, almost every problem will be characterized quickly and the fix
obvious to someone" or more compactly, "given enough eyeballs, all
\href{https://en.wikipedia.org/wiki/Software_bug}{bugs} are shallow"
\cite{raymond_cathedral_1999}. Given the size of the neuroimaging research community,
there are a large number of potential beta-testers and co-developers.
However, it has been difficult for researchers to collaborate on the
same code base due to slight differences in acquisition protocols,
storage conventions, researcher preferences, and study requirements.

@@

The abstraction of data and repositories in Arcana enables the same
workflow implementation to be applied to datasets stored in BIDS format
or XNAT repositories. A single code-base can therefore be containerized
into BIDS apps or XNAT pipelines without adaptation, helping to form a
bridge between the two communities of users and developers. Alternative
data storage systems \cite{,scott_coins:_2011,das_loris:_2012,book_neuroinformatics_2013}, can be integrated into Arcana by overriding a small number of
methods from the Repository abstract base class. Repository modules
could also be created for data portals such as \emph{DataLad} \cite{yaroslav_halchenko_datalad/datalad_2018} in order to take advantage of the range of platforms they support.
Implementing analyses in Arcana therefore enables researchers and
research groups to easily migrate their workflows between storage
platforms, and not risk being locked in to a particular technology.

@@

Ensuring that the consistent versions of external tools are used
throughout the analysis is important to avoid introducing biases due to
algorithm updates. In systems with environment modules \cite{furlani_modules:_1991}
installed, Arcana loads and unloads the required modules before and
after each node is executed. When running Arcana within a container,
environment modules and software versions can be installed inside the
container giving exact control over the versions used. To these ends, a
Docker container is available on Docker Hub,
(\href{http://hub.docker.com/r/monashbi/banana}{{http://hub.docker.com/r/monashbi/banana}}),
which can be used as a base for neuroimaging analysis containers.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{natbib}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{color}

@@

\lstset{
    basicstyle=\footnotesize,
    captionpos=b,                    % sets the caption-position to bottom
    language=Python,                 % the language of the code
    showstringspaces=false,          % underline spaces within strings only
    stringstyle=\color{stringcolour},     % string literal style
    commentstyle=\color{commentcolour},    % comment style
    keywordstyle=\color{keywordcolour}\bfseries,       % keyword style
}

@@

\author{
	Thomas G. Close\affmark[1,2]*\and
	Phillip G. D. Ward\affmark[1,3,4] \and
	Francesco Sforazzini\affmark[1] \and
	Wojtek Goscinski\affmark[6] \and
	Zhaolin Chen\affmark[1,5] \and
	Gary F. Egan\affmark[1,3,4]
}

@@

Despite the availability of well-established neuroimaging analysis
packages \citep{cox_afni:_1996,smith_advances_2004,friston_statistical_2007,tournier_mrtrix:_2012}, the arcana of neuroimaging analysis is substantial due to
the range of available tools, tuneable parameters, and imaging sequences
involved \citep{cusack_automatic_2015}. The distribution of complete
``end-to-end'' workflows, from acquired data to publication results, is
necessary for routine reproduction because of the effort required to
accurately exactly reimplement such analyses \citep{kennedy_neuroimaging_2018}. It is also
difficult to adapt existing workflows to new studies without detailed
knowledge of their design. Therefore, flexible, portable and complete
workflows are important to promote reproduction and code reuse in
neuroimaging research.

A barrier to designing portable and complete workflows is the
heterogeneity of data storage conventions \citep{marcus_extensible_2007,das_loris:_2012,gorgolewski_brain_2016}.
To address this, the emerging Brain
Imaging Data Standard (BIDS) \citep{gorgolewski_brain_2016} provides a way
to standardise the storage of neuroimaging data in file system
directories. BIDS specifies strict file and directory naming
conventions, which facilitate the design of portable \emph{BIDS Apps}
(https://bids-apps.neuroimaging.io). However, for research groups with
sufficient informatics support, software-managed repositories \citep{marcus_extensible_2007,das_loris:_2012} can provide additional features, such as
flexible access-control and automated pipelines. Therefore, for
published workflows, the choice of repository should be transparent in
order to maximise their audience.

While neuroimaging analyses are generally amenable to standardisation
\citep{kennedy_neuroimaging_2018}, minor modifications are often required to accommodate
idiosyncrasies of the acquisition protocols in use at different sites
\citep{esteban_fmriprep:_2018}. Therefore, workflows may require conditional
logic in construction or execution to be portable. Nipype is a flexible
Python framework for neuroimaging analysis in which workflows are
constructed programmatically in Python \citep{gorgolewski_nipype:_2011}.
Programmatic construction allows for rich control-flow logic that is not
readily available in alternative workflow frameworks \citep{cusack_automatic_2015,achterberg_fastr:_2016,amstutz_common_2016}, and has been used
to implement workflows that are robust to differences in fMRI protocols
across a large number of sites \citep{esteban_fmriprep:_2018}.

The trend towards large multi-site and multi-contrast datasets collected
over a number of years \citep{van_essen_human_2012,thompson_enigma_2014,sudlow_uk_2015}
presents additional challenges to workflow design.
Analysis packages are constantly being developed and improved, so the
state-of-the-art workflow for a particular analysis can change over
time. Therefore, it is challenging to ensure workflows are applied
consistently over the course of long studies \citep{cusack_automatic_2015}.

While analysis workflows for different contrasts and modalities are
typically implemented independently, they can share common processing
steps (e.g. non-linear registration to standard space, surface
parcellation) and their outputs may need to be integrated to produce
publication results. For large scale studies, which are typically
processed on the cloud or high-performance computing (HPC) clusters,
rerunning common segments can lead to significant increases in
computation time and therefore project cost. In addition, duplication of
processing segments increases time for manual quality control (QC),
making the reuse of intermediate derivatives a practical requirement for
some large studies \citep{schreiber_using_2018}.

@@

The utility of the Arcana framework is demonstrated by the
implementation of analysis suites for T1, T2* and diffusion weighted MRI
(dMRI) data and the application of dMRI tractogram \citep{tournier_mrtrix:_2012} and
composite vein image \citep{ward_combining_2018} workflows to data collected from a
healthy subject.

@@

\begin{figure}
\lstinputlisting[language=Python, firstline=2, lastline=58]{listings/example_study.py}
\caption{Example data and parameter specifications. The data
specification specifies two ``acquired'' file sets, `one' and `ten' and
ten derived file sets that can be derived from them, at least
indirectly. Each derived data spec, specifies the name of the pipeline
constructor that creates the pipeline that derives them. Parameter
specifications specify a name and default value for free parameters of
the Study class}
\label{list:data_spec}
\end{figure}

@@

Similar to data specifications, parameter specifications are included in
the Study class by providing a list of \emph{ParameterSpec} objects to
the \emph{add\_param\_specs} class attribute. ParameterSpec objects
are initialised with a name and default value. Special parameters that
specify a qualitative change in the analysis, for example using ANTs
registration \citep{avants_reproducible_2011} instead of FSL registration
\citep{smith_advances_2004}, are specified by the \emph{SwitchSpec} subtype.
SwitchSpecs take a name, default value and a list of accepted values at
initialisation.

@@

\begin{figure}
\lstinputlisting[language=Python, firstline=59]{listings/example_study.py}
\caption{Example pipeline constructor method. Pipelines are
created using the \emph{create\_pipeline} method of the Study class.
Pipeline objects are thin wrappers around Nipype workflows to allow for
namespace management}
\label{list:pipeline_constructor}
\end{figure}

@@

Any external software packages required by a node should be referenced
in the \emph{requirements} keyword argument as a list of
\emph{Requirement} objects when the node added to the pipeline. These
requirements can be used by the Processor to configure environment
modules \citep{furlani_modules:_1991} if present, to ensure the required software
packages are available. Similarly, the expected memory requirements and
wall time for the node execution should be provided as keyword args,
\emph{memory}, and \emph{wall\_time}, so they can be inserted into job
scripts submitted to schedulers by HPC Processors.

@@

Instead of setting the class attributes directly, the data and parameter
specifications are set by the metaclas\emph{s}, in order to combine with
specifications from base classes. The combined data and parameter
specifications are constructed by visiting the class' bases in reverse
method resolution order (MRO) and adding specifications from their
\emph{add\_data\_specs}, \emph{add\_param\_specs} attributes,
overriding previously added specifications with matching names. Note
that in this scheme, specifications can only be appended or overridden
but not removed by Study subclasses so as not to break workflows
inherited from base classes.

@@

All MultiStudy classes must inherit from the Multi\emph{Study} base
class and be created by the Multi\emph{StudyMetaClass} metaclass or
subclasses thereof. As in the case of subclassing the standard Study
class, additional data and parameter specifications can be added to the
class via \emph{add\_data\_specs} and \emph{add\_param\_specs}
respectively for additional analysis not included in the sub-studies.

@@

In Arcana, repository access is encapsulated within modular
\emph{Repository} objects to enable switching between different
repositories and repository types at analysis ``application time''
(Figure 2). There are currently three supported repository types, XNAT
\citep{marcus_extensible_2007}, BIDS \citep{gorgolewski_brain_2016} and a simple
``directory'' format, which are encapsulated by \emph{XnatRepository},
\emph{BidsRepository}, and \emph{DirectoryRepository} classes
respectively.

@@

\begin{table}
\caption{Storage locations of derived data for each repository
type. Derivatives are stored in separate namespaces for each Study
instance to enable multiple analyses on the same datasets with different
parameterisations. Where `...' is the location of the
directory or MrSession that holds the derivatives, \emph{subj} = subject ID, \emph{vis} = visit ID,
\emph{study} = name of the Study instance, and \emph{pl} = name
of pipeline.}
\label{tbl:repo-structure} 
\begin{tabular}{llll}
\hline\noalign{\smallskip}
\textbf{Type} & \textbf{Plain-directory} & \textbf{BIDS} & \textbf{XNAT}  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\textbf{Derivatives} & /\emph{subj}/\emph{vis}/\emph{study} & /derivatives/\emph{study}/\emph{subj}/\emph{vis} & /\emph{subj}/\emph{vis}\_\emph{study}\\
\textbf{Fields} & \D/\_\_fields\_\_.json & \D/\_\_fields\_\_.json & MrSession XML \\ 
\textbf{Provenance} & \D/\_\_prov\_\_/\emph{pl}.json & \D/\_\_prov\_\_/\emph{pl}.json & \D/\_\_prov\_\_/\emph{pl}.json \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

@@

\begin{table}
\caption{Abstract methods in the base Repository class that
need to be implemented by platform-specific sub-classes.}
\label{tbl:abstract-repo-methods} 
\begin{tabular}{ll}
\hline\noalign{\smallskip}
\textbf{Method name} & \textbf{Function} \\
\noalign{\smallskip}\hline\noalign{\smallskip}
find\_data() & Queries repository to find all existing filesets and
fields \\
get\_fileset(Fileset) & Caches file-set (if necessary) and returns the file path. \\
get\_field(Field) & Retrieves and returns the value of the field from
the repository \\
put\_fileset(Fileset) & Inserts file-set into the repository and updates cache. \\
put\_field(Field) & Inserts the value of the Field into the repository \\
connect() & Opens a connection to the repository (optional). \\
disconnect() & Closes connection to repository (optional)\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

@@

Software versions are managed by the processor if environment modules
\citep{furlani_modules:_1991} are installed on the system. These are loaded and
unloaded before the interface of a workflow node is run based on the
requirements specified for the node during construction of the pipeline
(see \emph{Pipeline constructors}. Maps from requirement names to
module names in the processing environment can be passed to the
Processor at initialisation. For plugins that submit jobs to HPC
schedulers (e.g. SLURM, SGE), node memory and wall-time requirements are
inserted into jobscripts where required.

@@

The Arcana framework is distributed as a publicly available software
package via GitHub
(\href{http://github.com/MonashBI/arcana}{{http://github.com/MonashBI/arcana}})
and the Python Package Index (PyPI)
(\href{https://pypi.org/project/arcana/}{{http://pypi.org/project/arcana/}}).
Study classes for T1, T2* and diffusion weighted MRI data have been
implemented as part of the \emph{Biomedical imAgiNg ANAlysis (Banana)} package \\
(\href{https://github.com/MonashBI/banana}{{http://github.com/MonashBI/banana}} and
\href{http://pypi.org./project/banana}{{http://pypi.org./project/banana}}).
All three classes, \emph{T1wStudy}, \emph{T2starwStudy} and
\emph{DmriStudy}, inherit generic image analysis methods, such as
registration and brain extraction, from the base class \emph{MRIStudy}.
The T1Study and T2starStudy classes are aggregated into the
\emph{T2starwT1wStudy} MultiStudy class to combine analysis of T2* and
T1-weighted contrasts.

The DmriStudy class implements diffusion tensor metrics, FA and ADC, as
well as global tractography using streamlines tracking from the MRtrix
toolbox \citep{tournier_improved_2010,tournier_mrtrix:_2012} (Figure 9).

@@

The \emph{T2starwT1wStudy} class implements an algorithm to generate
\emph{composite vein images} from the combination of QSM and SWI
contrasts derived from the T2*-weighted acquisition with an vein atlas
registered to the T1-weighted acquisition \citep{ward_combining_2018} (Figure 10).

@@

To test the analysis workflows implemented in ArcanaPaper, a healthy
volunteer was scanned using a 3T Siemens Skyra with a 32-channel head
and neck coil. The protocol was a T1-weighted MPRAGE (1mm contiguous,
matrix size 256x240x192, FOV 256x240x192, TE = 2.13ms, TR, = 2300ms, TI
= 900ms, bandwidth=230Hz/pixel), GRE (1.8 mm contiguous, matrix size
256x232x72, FOV 230x208x130, TE = 20ms, TR, = 30ms,
bandwidth=120Hz/pixel), diffusion MRI (1.2 mm contiguous, matrix size
110x100x60, FOV 256x240x192, TE = 95ms, TR, = 8200ms, 33 diffusion
directions with b=1500 mm2/s and 3 b=0, bandwidth=781Hz/pixel)
(\textbf{cite Lifespan paper ??}). Methods for Figure 11, 12, and 13 in
the ArcanaPaper class were applied to the acquired dataset to using the
\emph{SlurmProcessor} to submit jobs to the the MASSIVE HPC cluster
\citep{goscinski_multi-modal_2014} in order to generate the figures.

Figure \ref{fig:veins} displays composite vein images and vein masks for the healthy
volunteer along with the SWI and QSM intermediate derivatives and the
input vein atlas. The derived vein images are comparable to those
generated by the original implementation \citep{ward_combining_2018}. The complete
workflow took ?? min to complete over ?? compute cores.

@@

Software frameworks \citep{yacoub_pattern-oriented_2004} have been successful in
improving code quality and efficiency of development in a variety of
contexts \citep{moore_professional_2008,white_hadoop:_2012,abadi_tensorflow:_2016}. By
factoring out common elements, only features that are specific to the
given application need to be implemented by the analysis designer, and
the common elements become battle hardened through repeated use. Arcana
handles many of the menial tasks involved with workflow implementation,
such as data retrieval and storage, format conversions, and provenance,
reducing the time and effort required to implement robust workflows from
acquired data to publication results.

An oft-repeated mantra in the open-source software movement dubbed
Linus' Law is that "given a large enough beta-tester and co-developer
base, almost every problem will be characterized quickly and the fix
obvious to someone" or more compactly, "given enough eyeballs, all
\href{https://en.wikipedia.org/wiki/Software_bug}{bugs} are shallow"
\citep{raymond_cathedral_1999}. Given the size of the neuroimaging research community,
there are a large number of potential beta-testers and co-developers.
However, it has been difficult for researchers to collaborate on the
same code base due to slight differences in acquisition protocols,
storage conventions, researcher preferences, and study requirements.

@@

The abstraction of data and repositories in Arcana enables the same
workflow implementation to be applied to datasets stored in BIDS format
or XNAT repositories. A single code-base can therefore be containerized
into BIDS apps or XNAT pipelines without adaptation, helping to form a
bridge between the two communities of users and developers. Alternative
data storage systems \citep{,scott_coins:_2011,das_loris:_2012,book_neuroinformatics_2013}, can be integrated into Arcana by overriding a small number of
methods from the Repository abstract base class. Repository modules
could also be created for data portals such as \emph{DataLad} \citep{yaroslav_halchenko_datalad/datalad_2018} in order to take advantage of the range of platforms they support.
Implementing analyses in Arcana therefore enables researchers and
research groups to easily migrate their workflows between storage
platforms, and not risk being locked in to a particular technology.

@@

Ensuring that the consistent versions of external tools are used
throughout the analysis is important to avoid introducing biases due to
algorithm updates. In systems with environment modules \citep{furlani_modules:_1991}
installed, Arcana loads and unloads the required modules before and
after each node is executed. When running Arcana within a container,
environment modules and software versions can be installed inside the
container giving exact control over the versions used. To these ends, a
Docker container is available on Docker Hub,
(\href{http://hub.docker.com/r/monashbi/banana}{{http://hub.docker.com/r/monashbi/banana}}),
which can be used as a base for neuroimaging analysis containers.
